### 2021-10-31 introduction.tex
Introduction added

\section{Introduction}

Large-scale web services and decentralized applications often rely on
geo-distributed state replication to meet their latency and availability
needs. An application's state is replicated asynchronously in such a
setting, meaning that operations are independently executed at each
replica, and updates are applied asynchronously at other replicas after a
possible network delay. Asynchronous execution complicates programming and
reasoning about distributed applications as it induces the possibility of
conflicting updates leading to non-convergence and application integrity
violations. Two basic approaches have been proposed to address this
problem. The first approach is to selectively strengthen the system
consistency to pre-empt the conflicting updates. The second is to redefine
the application state in terms of \emph{Replicated Data Types} (RDTs) that
are specially engineered to handle conflicting updates. Strengthening
system consistency necessarily entails inter-replica coordination,
therefore applications prefer to utilize RDTs as much as possible,
resorting to consistency strengthening only when it is necessary to maintain
application integrity. The common design principle guiding the development
of replicated data types is commutativity. The idea is that if the
replicated state is only updated by commutative operations, then updates
can be applied in any order and the replica states are still guaranteed to
converge. Indeed, there exist common usecases in web applications that can
be implemented using Commutative Replicated Data Types
(CRDTs)~\cite{crdts}. For instance, a video view counter on a streaming
application such as YouTube can be implemented as a Counter CRDT that
supports commutative increments.

In general, however, commutativity is not a common occurrence among data
types. Most common data type definitions come with at least a pair of
operations that do not commute. For instance an \C{insert} and a \C{remove}
operations on a set do not commute if both are for the same element.
Likewise two \C{insert} operations for the same position in a list do not
commute. Such non-commutative operations translate to conflicting updates
in a distributed setting. For instance, concurrent \C{insert $k$} and
\C{remove $k$} operations for an element $k$ in a replicated set are in
conflict, and it is not clear what their cumulative result should
be\footnote{Here, the \C{insert} $i$ is assumed to have witnessed a
  \C{remove} $r'$ that is distinct from the earlier \C{remove} $r$. Both
  $i$ and $r'$ are concurrent to $r$, hence their effects are not visible
  during $r$'s execution.}. To use such a non-commutative data type in a
replicated setting, creative re-engineering of its internal representation
and algorithms is needed to turn it into a bonafide CRDT with a sensible
distributed semantics. In the case of set data type, that could involve
changing the internal representation to maintain two sets -- a set
containing the current elements and a \emph{tombstone} set that contains
every element that has ever been deleted from the set. A \C{remove
$k$} operation removes $k$ from the element set and adds it to the
tombstone set, and an \C{insert $k$} operation adds $k$ to the element set
only if it is not present in the tombstone set (i.e., only if $k$ was not
previously deleted from the set). This results in a CRDT set called
\emph{two-phase set} where an element once removed cannot be re-added.

A more complex re-engineering of the set data type could result in a set
CRDT with a different (and arguably more useful) semantics.
Sec.~\ref{sec:motivation} describes a \emph{remove-wins} CRDT set, which
allows elements to be readded, but which prioritizes \C{remove}s over
concurrent \C{insert}s. This transformation requires vector
clocks~\cite{vectorclock} to keep track of casual ordering among
\C{insert}s and \C{remove}s. In fact most CRDTs have carefully-crafted
implementations involving non-trivial technical devices like vector
clocks. For instance, Replicated Growable Array (RGA) -- a CRDT for
collaborative editing~\cite{rga}, makes use of a variant of Lamport
timestamps~\cite{lamportts}; TreeDoc, another collaborative-editing CRDT,
makes use of \emph{dense linear orders}~\cite{treedoc}; and JSON CRDT -- a
CRDT variant of the JSON storage format, implements transactional semantics
and causal dependency sets~\cite{jsoncrdt}. Such complexity makes it very
hard to reason about basic correctness properties of CRDTs, such as
convergence (formally known as \emph{strong eventual
consistency}~\cite{crdts}).  Indeed, formal verification of strong eventual
consistency for a subset of the aforementioned CRDTs required considerable
effort on behalf of multiple authors who are experts in the
field~\cite{kleppmann2017}. The need for such advanced implementation and
reasoning techniques makes the development of new CRDTs a research exercise
to be undertaken only by experts.

% hinders the widespread adoption of replicated data
% types in software development.
In this paper we propose an alternative approach to replicated data types
that addresses the problems discussed above. Our approach differs from the
CRDT approach in two fundamental ways. Firstly, we adopt a state-centric
model of replication with \emph{mergeability} of states as the defining
characteristic in contrast to CRDTs' operation-centric model characterized
by commutativity. Unlike commutativity, mergeability does not require a
full-scale re-engineering of the data type definition. Instead,
a data type only has to define a merge semantics in form a \C{merge}
function to reconcile concurrent versions of the type in presence of their
common ancestor version. Notably, \C{merge} function is left completely
unconstrained even as convergence is guaranteed on the corresponding
\emph{mergeable replicated data type} (MRDT)~\cite{mrdt}. Such strong
guarantees are possible due to the simplicity of state-based reasoning
relative to the operation- or trace-based one. Note that state-centric
model of replication has also been adapted to CRDTs, but such state-based
CRDTs require the replicated state to be organized as a lattice to
guarantee convergence -- a restriction we eliminate in our setting.

The second distinguishing aspect of our approach is the utilization of a
distributed runtime that orchestrates \emph{well-formed} distributed
executions that are guaranteed to converge. Note that it is easy to
construct a trivial runtime that guarantees convergence of all executions.
Such a runtime would make an extensive use of inter-replica coordination to
synchronize the execution of every operation and thus induce a
sequentially-consistent behavior. This would however defeat the purpose of
state replication as the latency incurred for synchronized execution of an
operation is prohibitively high, and the availability of the system during
the execution is correspondingly low. It is therefore important for a
distributed runtime of RDTs to not synchronize the execution of RDT
operations. Orchestrating a well-formed convergent execution without
impacting latency and availability may seem impossible due to the CAP
theorem~\cite{cap}, but it is not the case in fact. While CAP theorem does
rule out linearizability, it does not prohibit enforcing weaker consistency
constraints on distributed executions. In fact many CRDTs, including the
aforementioned \emph{remove-wins} set, implicitly assume \emph{causal
consistency}, which is weaker than linearizability but is stronger than the
default \emph{eventual consistency}. While causal consistency alone cannot
guarantee convergence, it sufficiently constrains the distributed execution
for the CRDT implementations to guarantee convergence. In this respect our
  approach differs in the sense that (a). It identifies a novel set of
  constraints on distributed execution that guarantee convergence of an
  MRDT \emph{regardless} of its merge semantics, and (b). It localizes
  those constraints spatially and temporally into the runtime such that
  per-operation latency and system-wide availability remain unaffected.
Sec.~\ref{sec:semantics} explains our runtime in detail.

\paragraph{Contributions} The contributions of this paper are summarized
below:
\begin{itemize}
  \item We present a model of state replication for \emph{mergeable} data
  types, i.e., data types equipped with a \C{merge} function, and identify a
  structural well-formedness condition over distributed executions that is
  sufficient to guarantee their convergence. This is in contrast to
  existing models of replicated data types, where each type is obligated to
  prove its convergence.

  \item We formalize a distributed runtime called \quark that orchestrates
  well-formed distributed executions. We prove the correctness of \quark,
  and also demonstrate its \emph{progress} property, namely that no
  distributed execution would ever get ``stuck''. Besides the manual proofs
  for the theorems, we also provide the full formalization of the \quark
  system and its properties in Alloy~\cite{alloy} and Ivy~\cite{ivy}. The
  Ivy formalization of \quark has been automatically verified with help of
  Z3~\cite{z3}. To the best of our knowledge, this is the first time an SMT
  solver was used to verify the \emph{meta-theory} of a formal setup.

  \item We realize an implementation of \quark as a light-weight shim layer
  on top of Scylla -- an off-the-shelf distributed key-value
  store~\cite{scylla}. An evaluation over several mergeable replicated data
  types (MRDTs) demonstrate that \quark has \emph{no} effect on the latency
  and availability of the system. To demonstrate the trade-offs not
  captured by latency and availability, we invent a new metric called
  \emph{staleness} and show that our approach does affect it negatively, albeit
  within the tolerable limits.
\end{itemize}