### 2021-11-12 motivation.tex
Motivation in progress

\section{Motivation}
\label{sec:motivation}

In this section we motivate our runtime-assisted approach to
convergent RDTs with help of a set data structure.
% demonstrates the need for run-time intervention to ensure convergence.

\subsection{Add-Wins Set RDT}

Fig.~\ref{fig:ocaml-set} shows a simplified interface of a set data
type in OCaml. The interface hides a reference to a set (\C{Set.t}),
which can be updated in-place via \C{insert} and \C{remove}
operations. As such \C{Set} is an ordinary data type primarily
designed for sequential execution. Under a concurrent execution with
an asynchronously replicated state, \C{Set} would exhibit anamolous
behaviours such as those in Fig.~\ref{fig:crdt-execs}.

Fig.~\ref{fig:crdt-execs-1} shows an anamolous execution with three
replicas -- $R_1$, $R_2$, and $R_3$, all of which start with a
singleton set containing the element $e$. A client connects to the
replica $R_3$ and executes a \C{remove $e$} operation, which is then
asynchronously propagated to other replicas. Some time after applying
$R_3$'s \C{remove} at $R_2$, another client connects to $R_2$ and
re-adds $e$ by issuing an \C{insert $e$} operation. Consequently, the
state at $R_2$ is again the singleton set $\{e\}$. Replica $R_3$
however receives $R_2$'s \C{insert} ahead of $R_1$'s \C{remove},
applies them in the same order, and ends up with an empty set. The
execution thus results in divergent replica states.

Note that if set \C{insert} and \C{remove} commute, then executing
them in different order at $R_2$ and $R_3$ would not have led to
divergence. As such, \C{Set} is not a CRDT due to the admittance of
non-commutative operations. Nonetheless \C{Set} can be transformed
into a CRDT by re-engineering its interface and operations. For
instance, the anomalous execution in Fig.~\ref{fig:crdt-execs-1} can
be pre-empted by ensuring that updates are only ever applied in the
causal order.  This can be done by extending \C{Set} with vector
clocks to keep track of the causal history of each operation. A set
\C{insert} (resp. \C{remove}) would now generate an \C{Insert} (resp.
\C{Remove}) \emph{effect} tagged with the vector clock of the origin
replica. Here, the vector clock simply records the sequence number of
last operation from each replica whose effect has been received and
applied at the current replica. When an effect is received at a
replica, it is buffered until the time all its causally-preceding
effects (as captured by the tagged vector clock) have already been
received and applied. This strategy would preempt the execution in
Fig.~\ref{fig:ctrdt-execs-1} by buffering $R_2$'s \C{insert} at $R_1$
until the causally-preceding \C{remove} of $R_1$ is received and
applied. An interface for such a \emph{causally-consistent} set RDT is
shown in Fig.~\ref{fig:cc-set}.

Unfortunately, \C{Set} data type of Fig.~\ref{fig:cc-set} still admits
divergent executions due to concurrent updates.
Fig.~\ref{fig:crdt-execs-2} describes one such execution. Here,
replicas $R_1$ and $R_2$ both start with a singleton set $\{e\}$.
Two distinct clients connect to $R_1$ and $R_2$ respectively and issue
two concurrent \C{remove $e$} operations. Later, another
client connects to $R_2$ and issues an \C{insert $e$} operation. The
effects of these operations are asynchronously applied at remote
replicas as shown in the figure, resulting in the divergent states at
$R_1$ and $R_2$.

Note that, unlike the execution in Fig.~\ref{fig:crdt-execs-1}, the
conflicting operations in Fig.~\ref{crdt-execs-2}, namely $R_1$'s
\C{remove} and $R_2$'s \C{insert}, are \emph{not} causally related,
hence their relative order is not determined by the sequential
specification of the data type. Forcing a causal relationship between
them requires synchronization between \C{insert}s and \C{remove}s,
which is expensive in an asynchronous distributed setting. It
therefore becomes inevitable to ascribe semantics to concurrent
executions to restore convergence. This is done by imposing an
\emph{arbitration order} among concurrent conflicting operations,
which are otherwise unordered. For example, in
Fig.~\ref{fig:crdt-execs-2}, we might let $R_2$'s \C{insert} override
$R_1$'s \C{remove} considering that \C{insert} is re-inserting an
element after a previous remove. Ordering concurrent removes ahead
of inserts uniformly on all replicas leads to an implementation of
\C{Set} RDT where concurrent (re-)insertions consistenty \emph{win}
over removals. Such an \emph{add-wins} set is useful, for instance, to
implement a shared shopping cart where two users can concurrently
remove an item from the cart, but if one of them re-adds it then the
item should be present in the final cart\footnote{This is in fact the
semantics of Amazon's shopping cart~\cite{dynamo}}.

Extending the \C{Set} implementation from Fig.~\ref{fig:cc-set} with
\emph{add-wins} semantics is however not trivial. The suggested
approach involves tracking element-wise causal dependencies between
the \C{remove}s and \C{insert}s by maintaining a vector clock
\emph{for each element} $e$ in the set~\cite{zawirski-thesis}. The
vector clock of $e$ records the sequence number of the last \C{insert
$e$} operation from each replica whose effect has been received and
applied at the current replica. The \C{Insert} and \C{Remove} effects
on $e$ will now be tagged with $e$'s vector clock. When an \C{Remove}
effect on $e$ is received at a replica, it is applied only if the
tagged vector clock is no less than $e$'s local vector clock, i.e.,
only if the arriving \C{Remove} has seen (i.e., causally succeeds) at
least those \C{insert $e$} operations the current replica is aware of.
Otherwise the effect is simply a no-op. This strategy would result in
convergent states despite the execution in Fig.~\ref{fig:crdt-execs-2}
as $R_2$'s \C{remove} effectively becomes a no-op at $R_1$ due to
there being at least one \C{insert} operation on $e$ that it hasn't
seen. The \C{Set} interface extended with element-wise vector clocks
is shown in Fig.~\ref{fig:rw-set}. The resultant set RDT is assumed to
be correct albeit a formal proof of convergence could not be found in
the literature.

The above exercise demonstrates the considerable ingenuity and effort
involved in deriving a convergent RDT out of such simple data type as
\C{Set}. Some of the effort can be mitigated by strengthening the
underlying system model insofar as it doesn't affect the latency and
availability of the application. For instance, a system that always
delivers messages in the causal order would automatically preempt the
execution in Fig.~\ref{fig:crdt-execs-1} without the need for
additional intervention on behalf of the developer. This is a
particularly attractive proposition considering that causal
consistency can be ``bolted on'' an existing implementation of an
eventually consistent system without weakening its
guarantees~\cite{bolton}. Unfortunately, such strengthening of the
system model would deliver no benefits to the developer if they still
have to reason about fine-grained causal dependencies to guarantee
convergence, such as in the case of Fig.~\ref{fig:crdt-execs-2}. As we
observed earlier, the execution in Fig.~\ref{fig:crdt-execs-2} seems
inevitable unless every pair of \C{insert} and \C{remove} operations
are synchronized, which, regrettably, is not a practical option. The
developer therefore seems to be stuck.

Fortunately, there is a way out of this problem and the solution
requires a switch in the perspective of replication from an
operation-centric view characterized by explicit effects to a
state-centric view characterized by explicit state merges. The
motivation for this switch is the key observation that it is possible
to induce an ordering over \emph{states} of replicas even as their
operations and effects remain unordered, as long as such states are
\emph{mergeable}. The ordering is enforced by synchronizing merges in
the background while the replica-local execution progresses
unhindered. Consequently, eventual convergence can be guaranteed
\emph{without} impacting the user-perceived latency of the operations.
Moreover, working with states instead of operations frees the latter
from having to conform to algebraic laws such as commutativity.  Put
together, this means that the clients of a replicated data type can
perform any action allowed by the sequential version of the type and
immediately see the results of their actions on the local replica
while being safe in the knowledge that replica states will eventually
converge.

To demonstrate the aforementioned intuitions we reconsider the
executions from Fig.~\ref{fig:crdt-execs}, this time in the
state-centric replication model. The cornerstone of the state-centric
model is a three-way \C{merge} function that merges concurrent
versions of the state in presence of their (lowest) common ancestor
version. In our running example, the state is a value of type
\C{Set.t}, hence \C{Set.merge} function would have the type signature:
\begin{center}
\C{Set.merge} : \C{Set.t} $\rightarrow$ \C{Set.t} $\rightarrow$ \C{Set.t}
$\rightarrow$ \C{Set.t}
\end{center}
The three arguments of \C{merge} correspond to the lowest common
ancestor (LCA) version and the two concurrent versions that
independently evolved from the ancestor version. The LCA is a causal
ancestor of concurrent versions, hence causal consistency is built
into the replication model. The result of set merge, intuitively, must
contain the common elements in the two concurrent versions along with
any newly added elements in either versions. Concretely:
\begin{ocaml}
let merge s s1 s2 = 
              (s1 $\cap$ s2) $\cup$ (s1 - s) $\cup$ (s2 - s)
\end{ocaml}
Equipped with the above set merge function, we can now consider the
equivalent execution of Fig.~\ref{fig:crdt-execs-1} in state-centric
model. The execution is shown in Fig.~\ref{fig:mrdt-execs-1}. The
initial version on all three replicas ($v_0$) is the singleton set
$\{e\}$. Applying operations to replicas creates new versions, e.g.,
$v_1$ on $R_3$. Changes can be propagated by merging versions, e.g.,
version $v_2$ on $R_2$ is a result of merging $v_0$ and $v_1$ in
presence of their LCA $v_0$\footnote{Versions $v_0$ and $v_1$ are not
  concurrent as the former is an ancestor of the latter. Merging $v_1$
  into $v_0$ is nonetheless possible as $v_1$ is ahead of $v_0$ in
  causal order. In Git parlance this is a \emph{fast forward} merge.}.
Likewise $v_4$ on $R_1$ is created by merging $v_3$ and $v_0$ in
presence of their LCA $v_0$. By the end of the execution, versions $v_4$
and $v_3$ at $R_1$ and $R_2$ (resp.) have witnessed the same set of
operations, hence are in agreement.

Convergence however is not an inherent virtue of the state-centric
replication model. Fig.~\ref{fig:mrdt-exec-2} shows the state-centric
analogue of the execution in Fig.~\ref{fig:crdt-exec-2}, which too
diverges.