### 2021-11-17 concrete-semantics.tex
Quark concrete semantics added

\section{Concrete Semantics}
\label{sec:concrete-sem}

\input{quark-semantics}