### 2021-11-20 related.tex
Quark --> Carmot. Related work.

\section{Related Work}
\label{sec:related}

The idea of state-based replicated data types with three-way merge
functions was first proposed by Kaki et. al. in~\cite{mrdt},
where authors demonstrate an approach for deriving merge functions
from first principles for various data types. However, the resultant
\emph{mergeable replicated data types} (MRDTs) were never proved to be
convergent. As we demonstrate for the \C{Set} data type in
Sec.~\ref{sec:motivation}, simple merge functions with intuitive merge
semantics can nonetheless lead to divergent executions. The \quark
preempts such executions to guarantee convergence on MRDTs.

Concurrent Revisions~\cite{BBL+10} proposes versioned state for
shared-memory concurrent programs in order to guarantee determinism.
The fork-join concurrency they consider is however too restrictive for
the asynchronous distributed setting where replicas do not have a
single coherent view of the system state.  Furthermore the
hierarchical organization of the concurrent actors with a designated
\emph{server} is inconsistent with the fully-decentralized model of
execution adopted by several collaborative applications.

State-centric replication was also explored in the context of
CRDTs~\cite{crdts}. However, such state-based CRDTs require the
replicated state to be organized as a lattice with the merge function
acting as as a least upper bound operator. We eliminate this
restriction in our setting with help of the \quark runtime. A thorough
comparison of our approach with the operation-based CRDTs can be found
in Secs.~\ref{sec:introduction} and~\ref{sec:motivation}.

Several verification techniques, program analyses, and tools have been
proposed to reason about the program behavior in a weakly-consistent
distributed setting~\cite{bailis-vldb, alvaro-calm,
gotsman-popl16,redblue-atc, redblue-osdi, ecinec}. These techniques
treat replicated storage as a black box with a fixed pre-defined
consistency models. The focus is on assigning appropriate consistency
levels to operations so as to preserve application integrity. Such an
approach results in assigning sequential consistency (SC) to all
operations since the next weaker consistency model -- causal
consistency, is insufficient to guarantee convergence (as
Sec.~\ref{sec:motivation} demonstrates). Conversely, \quark cannot
reason about application-level invariants, such as $\C{balance}\ge 0$
in an banking application. Thus both approaches confer complimentary
benefits on application developers.

Finally, the implementation of \quark bears resemblance to
Quelea~\cite{pldi15}. Both these systems are implemented as shim
layers on top of off-the-shelf distributed stores, and both offload
low-level concerns to the underlying store. However, both these
systems serve orthogonal purposes. Quelea enforces high-level
invariants on applications built with CRDTs. Operations are assumed to
be commutative and applications are assumed to be convergent. In
contrast, \quark enforces convergence of data types that are otherwise
not guaranteed to converge.